---
title: "R Course Project"
output:
  pdf_document: default
  html_notebook: default
---


# dominant crop types
- Corn
- Wheat
- Soybeans



# Different methods of fertilization
Based on the dataset we have, four main methods are used for applying fertilizer:
- Applied on living crop
- broadcast with no incorporation
- broadcast with incorporation
- injected



# packages
required packages are all listed here

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Ryerson/R")
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(readxl)
#require(foreign)
require(nnet)
require(ggplot2)
require(ggplot)
require(reshape2)
#require(dplyr)
#require(caTools)
require(lessR)
#require(stringr)
require(forcats)
require(caret)

```

# manure
input 

```{r , include=FALSE}
manure <- read_excel('AGRIModelDATA_MASTER_ORIGINAL_2June2021.xlsx', sheet = "Detail_Manure")
manure <- manure[-c(6,7)]
manure[c(1,2)] <- lapply(manure[c(1,2)], as.numeric)
manure[c(1,4,6,7,13)] <- lapply(manure[c(1,4,6,7,13)],as.factor)
manureApplication.table <- table(manure$Application)
manureApplication.table

manure <- manure[manure$Crop == "corn" | manure$Crop =="wheat" | manure$Crop == "soybeans", ]
manure <- droplevels(manure[!manure$Crop == "corn" | !manure$Crop =="wheat" | !manure$Crop == "soybeans",])

str(manure)
manure.croptable <- table(fct_infreq(manure$Crop))
manure.croptable
```



## Detail_Fertilizer

```{r, include=FALSE}
fertilizer <- read_excel('AGRIModelDATA_MASTER_ORIGINAL_2June2021.xlsx', sheet = "Detail_Fertilizer")
names(fertilizer)[11] <- "Fert_Date_End"
fertilizer[c(1,2)] <- lapply(fertilizer[c(1,2)], as.numeric)
fertilizer <- fertilizer[!is.na(fertilizer$Crop),]
fertilizer[c(1,4,6,9,16,17)] <- lapply(fertilizer[c(1,4,6,9,16,17)],as.factor)

```



```{r, results="hide"} 
FertilizedData <- fertilizer[!is.na(fertilizer$HasFertilizer), ]
#FertilizedData[FertilizedData$HasFertilizer == "Yes",]$HasFertilizer <- 1
#FertilizedData[FertilizedData$HasFertilizer == "No",]$HasFertilizer <- 0
FertilizedData$HasFertilizer <- as.factor(FertilizedData$HasFertilizer)

```





## Lostic regression





```{r}
Model1 <- Logit(HasFertilizer~ Crop, data= FertilizedData)
reg(HA~ Crop, data= FertilizedData)

```

```{r}
Model2 <- Logit(HasFertilizer~ HA, data= FertilizedData , prob_cut=0.8)
```


HasFertilizer vs. HA + Crop
```{r}
#multiple Logistic Regression model
model3 <- Logit(HasFertilizer~ HA + Crop, data= FertilizedData)
caret::varImp(model3)
```





## Multinomial logistic Regression Analysis
manure application method vs. crop type and area

For all three crop types the probability of using "broadcast with incorporation" reduces significantly for larger fields.

```{r}

with(manure, table(Crop, Application))
with(manure, do.call(rbind, tapply(HA, Application, function(x) c(M = mean(x), SD = sd(x)))))

manure$Application2 <- relevel(manure$Application, ref = "applied on living crop")
testManure <- multinom(Application2 ~ Crop + HA, data = manure)

summary(testManure)

zManure <- summary(testManure)$coefficients/summary(testManure)$standard.errors
zManure

# 2-tailed z test
pManure <- (1 - pnorm(abs(zManure), 0, 1)) * 2
pManure

## extract the coefficients from the model and exponentiate
exp(coef(testManure))

head(ppManure <- fitted(testManure))

dHAManure <- data.frame(Crop = rep(c("corn", "soybeans","wheat"), each = 166), HA = rep(c(0:165), 3))


## store the predicted probabilities for each value of Crop and HA
pp.HAManure <- cbind(dHAManure, predict(testManure, newdata = dHAManure, type = "probs", se = TRUE))

## calculate the mean probabilities within each level of Crop
by(pp.HAManure[, 3:5], pp.HAManure$Crop, colMeans)


## melt data set to long for ggplot2
lppManure <- melt(pp.HAManure, id.vars = c("Crop", "HA"), value.name = "probability")
head(lppManure)  # view first few rows

## plot predicted probabilities across HA values for each level of Crop
## faceted by Application type
ggplot(lppManure, aes(x = HA, y = probability, colour = Crop)) + geom_line() + facet_grid(variable ~   ., scales = "free")
  
```


## Fertilizer Analysis 

Fertilizer application method vs. crop type and area

```{r}

fertilizerApplication.table <- table(fertilizer$Application)
fertilizer <- fertilizer[fertilizer$Crop == "corn" | fertilizer$Crop =="wheat" | fertilizer$Crop == "soybeans", ]
fertilizer <- droplevels(fertilizer[!fertilizer$Crop == "corn" | !fertilizer$Crop =="wheat" | !fertilizer$Crop == "soybeans",])
fertilizer$Application <- as.factor(fertilizer$Application)
str(fertilizer)
fertilizer.croptable <- table(fct_infreq(fertilizer$Crop))
fertilizer.croptable


with(fertilizer, table(Crop, Application))
with(fertilizer, do.call(rbind, tapply(HA, Application, function(x) c(M = mean(x), SD = sd(x)))))



fertilizer$Application2 <- relevel(fertilizer$Application, ref = "applied on living crop")
testFertilizer <- multinom(Application2 ~ Crop + HA, data = fertilizer)

summary(testFertilizer)

zFertilizer <- summary(testFertilizer)$coefficients/summary(testFertilizer)$standard.errors
zFertilizer

# 2-tailed z test
pFertilizer <- (1 - pnorm(abs(zFertilizer), 0, 1)) * 2
pFertilizer

## extract the coefficients from the model and exponentiate
exp(coef(testFertilizer))

head(ppFertilizer <- fitted(testFertilizer))

dCropFertilizer <- data.frame(Crop = c("corn", "soybeans","wheat"), HA = mean(fertilizer$HA))

dHAFertilizer <- data.frame(Crop = rep(c("corn", "soybeans","wheat"), each = 166), HA = rep(c(0:165), 3))


## store the predicted probabilities for each value of Crop and HA
pp.HAFertilizer <- cbind(dHAFertilizer, predict(testFertilizer, newdata = dHAFertilizer, type = "probs", se = TRUE))

## calculate the mean probabilities within each level of Crop
by(pp.HAFertilizer[, 3:5], pp.HAFertilizer$Crop, colMeans)


## melt data set to long for ggplot2
lppFertilizer <- melt(pp.HAFertilizer, id.vars = c("Crop", "HA"), value.name = "probability")
head(lppFertilizer)  # view first few rows

## plot predicted probabilities across HA values for each level of Crop
## faceted by Application type
ggplot(lppFertilizer, aes(x = HA, y = probability, colour = Crop)) + geom_line() + facet_grid(variable ~   ., scales = "free")
```



